NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2020-05-15 Manuscript/manuscriptPGPE.tex
Update discussion about headgroup structures

Without further discussion about poorly performing force fields, we focus on more detailed analysis
of CHARMM36 simulations, which give the best overall agreement with experiments for
the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
not within the experimental error bars, it captures the essential differences between different headroups
with the exception of $\beta$-carbon order parameter of PC (Fig.~\ref{structures}),
which is too large when compared with PS or PE order parameter, or with experiments \cite{botan15}.
As already pointed out in our previous study \cite{antila19}, N-C$_\beta$-C$_\alpha$-O$_\alpha$
dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
in gauche 270$^\degree$ conformation, than for the PC headgroup.
Also C$_\beta$-C$_\alpha$-O$_\alpha$-P in PS has asymmetric tendency to be in this conformation.
In PG, O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other lipids)
is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
gauche 60$^\degree$ conformation.

>>>>

Without further discussion about poorly performing force fields, we focus on more detailed analysis
of CHARMM36 simulations, which captures the essential differences between
PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
%give the best overall agreement with experiments for
%the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
%studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
%not within the experimental error bars, it 
which is too negative when compared with PS or PE order parameter, or with experiments \cite{botan15}.
%As already pointed out in our previous study \cite{antila19},
Characteristic dihedral conformations in PS headgroup are asymmetric conformations
preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
%in , than for the PC headgroup.
C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other lipids)
is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
gauche 60$^\circ$ conformation. 
Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.
